### 2020-02-25 paper/introduction.tex

The Single-fluid Double-zone Thorium-based Molten Salt Reactor (SD-TMSR) with a thermal power of 2,250
MW$_{th}$ was proposed for the first time by ORNL as early as in the 1960s, which was called Molten salt breeding reactor (MSBR) \cite{robertson_conceptual_1971}.
The SD-TMSR is a graphite-moderated thermal-spectrum \gls{MSR} operating in Th/$^{233}$U
fuel cycle. In the SD-TMSR, the fissile and fertile elements are integrated
into a single salt. To improve the breeding ratio, the active core is divided into two zones:
the radius of the fuel channels in the outer zone is modified to be larger
than the radius of the fuel channels in the inner zone \cite{li_optimization_2018}.

@@

Cui \emph{et al.} analyzed the fuel transition from enriched $^{235}$U/Th and Pu/Th to $^{233}$U/Th for the MSBR-like design. Cui \emph{et al.} found that the fuel transition can be achieved by using: (1) enriched uranium with greater than 40\% enrichment, which rises the proliferation concerns; (2) Pu from LWR (burn-up of 60 GWd/t) spent nuclear fuel \cite{cui2017transition,cui2018possible}.

The research effort described in \cite{cui2017transition} and
\cite{cui2018possible} is most similar to the work
presented in this paper. However, a few major differences
are: (1) we used built-in truly continuous fuel depletion capabilities in
SERPENT-2 \cite{aufiero2013extended} while Cui \emph{et al.} employed an in-house tool named MSR
Reprocessing Sequence (MSR-RS) \cite{ZOU2015114} which couples with SCALE and employs a batch-wise approach (the burnup simulation stops at a given time and restarts with a new liquid fuel composition after removal of discarded materials and addition of fissile/fertile materials); (2) we investigated different initial loading cases (HALEU, Pu+HALEU, reactor-grade Pu, TRU and $^{233}$U) while Cui \emph{et al.} studied only uranium and Pu as the startup fissile materials; (3) we adopted the SD-TMSR core geometry optimized by Li \emph{et al.} \cite{li_optimization_2018} which is different from the MSBR-like model adopted by Cui \emph{et al.}; (4) Cui \emph{et al.} studied the high-enriched uranium (e$>$90) while we excluded it due to the unavailability of this material for use as an initial reactor fuel (non-proliferation issues) and because the performance of this material is expected to be similar to $^{233}$U; (5) Cui \emph{et al.} introduced two scenarios for thorium fuel cycle transition: a Breeding and Burning (B\&B) scenario and a Pre-breeding scenario while we adopting different approach by introducing two different feed mechanisms: thorium and non-thorium feed mechanism.

Comparing with Cui \emph{et al.}, we used different geometry, volumes, densities, and fuel salt compositions. In Cui's scenarios, the amount of Th is kept constant. $^{233}$U is fed only in (B\&B) scenario and extra fissile materials would have to be added into the core if the produced $^{233}$U is not enough to maintain the reactor criticality. In the Pre-breeding scenario, all the $^{233}$U produced from the decay of the extracted $^{233}$Pa stored outside the core until it reaches the required startup mass for a new reactor. Thus, the criticality of the core is maintained by refueling external fissile material. 
In current work, using the thorium feed mechanism, we simultaneously feed Th and all or part of produced $^{233}$U from the \texttt{Pa-decay tank} into the core. The excess $^{233}$U (if exist) is stored outside the core. In the non-thorium feed mechanism, we continuously inject external heavy metals (HALEU, Pu mixed with HALEU, reactor-grade Pu, or TRU) and part of produced $^{233}$U from the \texttt{Pa-decay tank}. Here we feed a part of produced $^{233}$U because in some cases (e.g., TRU) refueling only external heavy metals failed to maintain the core criticality (limitation is dm$_{total fuel}$ $\leq$ 0.1\%). In the non-thorium feed mechanism, the amount of Th decreases during reactor operation while it is constant in the thorium feed mechanism.

>>>>

The Single-fluid Double-zone Thorium-based Molten Salt Reactor (SD-TMSR) with 
a thermal power of 2,250 MW$_{th}$ was proposed for the first time by ORNL as 
early as in the 1960s, which was called Molten Salt Breeder Reactor (MSBR) 
\cite{robertson_conceptual_1971}.
The SD-TMSR is a graphite-moderated thermal-spectrum \gls{MSR} operating in Th/$^{233}$U
fuel cycle. In the SD-TMSR, the fissile and fertile elements are integrated
into a single salt. To improve the breeding ratio, the active core is divided into two zones:
the radius of the fuel channels in the outer zone is modified to be larger
than the radius of the fuel channels in the inner zone \cite{li_optimization_2018}.

@@

The research effort described in \cite{cui2017transition} and
\cite{cui2018possible} is most similar to the work
presented in this paper. Cui \emph{et al.} analyzed the fuel transition from 
enriched $^{235}$U/Th and Pu/Th to $^{233}$U/Th for the MSBR-like design. Cui 
\emph{et al.} found that the fuel transition can be achieved by using: (1) 
enriched uranium with greater than 40\% enrichment \cite{cui2018possible}; (2) 
Pu from LWR (burn-up of 60 GWd/t) spent nuclear fuel \cite{cui2017transition, 
cui2018possible}.

However, a few major differences follow: 
(1) Cui \emph{et al.} employed an in-house tool named MSR Reprocessing 
Sequence (MSR-RS) \cite{ZOU2015114} which couples with SCALE and employs a 
batch-wise approach while we used built-in truly continuous fuel depletion 
capabilities in SERPENT-2 \cite{aufiero2013extended}; 
(2) Cui \emph{et al.} studied a high-enriched uranium ($>$40\% of $^{235}$U) 
and plutonium partitioned from light water reactors spent fuel as the startup 
fissile materials while we considered different loading cases: HALEU, Pu 
mixed with HALEU, reactor-grade plutonium, TRU and $^{233}$U;
(3) we adopted the SD-TMSR core geometry optimized by Li \emph{et al.} 
\cite{li_optimization_2018} which is different from the MSBR-like geometry by 
Mathieu \emph{et al.} \cite{mathieu_possible_2009} adopted by Cui \emph{et 
al.}; 
(4) Cui \emph{et al.} introduced two scenarios for thorium fuel cycle 
transition: a Breeding and Burning (B\&B) scenario and a Pre-breeding scenario 
while we adopting different approach by introducing two different feed 
mechanisms: thorium and non-thorium.

Comparing with Cui \emph{et al.}, we used different reactor design (geometry, 
volumes, densities, fuel salt compositions) and computation method (continuous 
online reprocessing instead of batch-wise approach). In Cui's scenarios, the 
amount of thorium is kept constant. $^{233}$U is fed only in (B\&B) scenario 
and extra fissile materials would have to be added into the core if the 
produced $^{233}$U is not enough to maintain the reactor criticality. In the 
Pre-breeding scenario, all the $^{233}$U produced from the decay of the 
extracted $^{233}$Pa stored outside the core until it reaches the required 
startup mass for a new reactor. Thus, the criticality of the core is 
maintained by refueling external fissile material. In the current work, using 
the thorium feed mechanism, we simultaneously feed thorium and all or fraction 
of produced $^{233}$U from the \texttt{Pa-decay tank} into the core. The 
excess $^{233}$U (if exists) is stored outside the core. In the non-thorium 
feed mechanism, we continuously inject external heavy metals (HALEU, Pu mixed 
with HALEU, reactor-grade Pu, or TRU) and fraction of produced $^{233}$U from 
the \texttt{Pa-decay tank}. For the non-thorium mechanism, we feed a part of 
produced $^{233}$U because in some cases (e.g., TRU) refueling only external 
heavy metals failed to maintain the core criticality (limitation is 
dm$_{salt}$ $\leq$ 0.1\%).

### 2020-02-25 paper/methods.tex

As shown in Figure~\ref{fig:flow}, after launching the input file, SERPENT solves the Bateman equation using an advanced 
matrix exponential solution based on the Chebyshev Rational Approximation 
Method \cite{isotalo2016improving}. 
Then, the system extracts gaseous \gls{FPs} and other materials 
(non-dissolved metals, lanthanides, and soluble metals except Pa) with 
a suitable removal rate\footnote{The extraction rate depends on the type of 
poison and its impact on the neutron
economy.}. This is done by setting the 
flow rate of gaseous \gls{FPs} and other materials from the fuel to the 
\texttt{FP-waste tank}\footnote{An external tank used to store the gaseous 
\gls{FPs} and the other materials (non-dissolved metals, lanthanides, and 
soluble metals except protactinium).}. Specifically, Pa is removed 
from the fuel with a certain flow rate into the 
\texttt{Pa-decay tank} to decay and produce $^{233}$U \footnote{The 
$^{233}$Pa is removed and left to decay into $^{233}$U with $\tau_{1/2}$ 
$\approx$ $27$ $d$.}. The produced $^{233}$U is used as a fresh fissile fuel 
and the residual $^{233}$U (if exist) is the net production of $^{233}$U that required to start up a new SD-TMSR and achieve the transition to thorium fuel cycle. The MSR burnup 
routine provided by SERPENT-2 allows changes to the mass flow rates ($mflow$) of the 
isotopes during reactor operation \cite{aufiero2013extended}. Specifically, the SERPENT user should determine the mass flow rate ($mflow$), which is the rate by 
which elements or nuclides are transferred between materials. After that, the 
flow rates must be connected to materials with a reprocessing scheme. 
The final step is to link the reprocessing schemes to depletion histories. In 
the present work, we adjust the transfer rates of fresh fuel to maintain core 
criticality and to keep fuel salt mass constant during burnup. The continuous feed constant calculation procedures are summarized as follows:
\begin{enumerate}
	\item The simulation starts without injecting refueling materials (i.e. only removing FPs and Pa).
	\item After the first depletion calculation step, we check the total mass density of FPs and Pa in the \texttt{FP-waste tank} and \texttt{Pa-decay tank}, respectively.
	\item A simple calculation yields the amount of heavy metal that must be added during this cycle:
	\subitem 3a. \textbf{Thorium feed mechanism}: mass of injected $^{233}$U $\approx$ mass of extracted Pa and mass of injected Th $\approx$ mass of extracted FPs to maintain the total fuel mass constant.
	\subitem 3b. \textbf{Non-thorium feed mechanism}: mass of injected $^{233}$U $\approx$ mass of extracted Pa and mass of injected fuel (i.e., HALEU, Pu mixed with HALEU, reactor-grade Pu, or TRU) $\approx$ mass of extracted FPs to maintain the total fuel mass constant.
	\item Dividing this mass by time and inventory of refueling material gives the corresponding feed constant (1/s).
\end{enumerate}
In both feed mechanisms (thorium and non-thorium) we continuously added $^{233}$U and the mass of injected $^{233}$U is equal to the mass of extracted Pa ($^{233}$Pa decays to $^{233}$U after $\approx$ 27 d). To keep the total fuel mass almost constant, for \textbf{thorium feed mechanism}, the mass of injected Th must be equal to the mass of extracted FPs and for \textbf{non-thorium feed mechanism}, the mass of injected fuel (i.e., HALEU, Pu mixed with HALEU, reactor-grade Pu, or TRU) must be equal to the mass of extracted FPs.

>>>>

As shown in Figure~\ref{fig:flow}, after launching the input file, SERPENT solves the Bateman equation using an advanced 
matrix exponential solution based on the Chebyshev Rational Approximation 
Method \cite{isotalo2016improving}. 
Then, the system extracts gaseous \gls{FPs} and other materials 
(non-dissolved metals, lanthanides, and soluble metals except Pa) with 
a suitable removal rate\footnote{The extraction rate depends on the type of 
poison and its impact on the neutron
economy.}. This is done by setting the 
flow rate of gaseous \gls{FPs} and other materials from the fuel to the 
\texttt{FP-waste tank}\footnote{An external tank used to store the gaseous 
\gls{FPs} and the other materials (non-dissolved metals, lanthanides, and 
soluble metals except protactinium).}. Specifically, Pa is removed 
from the fuel with a certain flow rate into the 
\texttt{Pa-decay tank} to decay and produce $^{233}$U \footnote{The 
$^{233}$Pa is removed and left to decay into $^{233}$U with $\tau_{1/2}$ 
$\approx$ $27$ $d$.}. The produced $^{233}$U is used as a fresh fissile fuel 
and the residual $^{233}$U (if exist) is the net production of $^{233}$U that required to start up a new SD-TMSR and achieve the transition to thorium fuel cycle. The MSR burnup 
routine provided by SERPENT-2 allows changes to the mass flow rates ($mflow$) of the 
isotopes during reactor operation \cite{aufiero2013extended}. Specifically, the SERPENT user should determine the mass flow rate ($mflow$), which is the rate by 
which elements or nuclides are transferred between materials. After that, the 
flow rates must be connected to materials with a reprocessing scheme. 
The final step is to link the reprocessing schemes to depletion histories. In 
the present work, we adjust the transfer rates of fresh fuel to maintain core 
criticality and to keep fuel salt mass constant during burnup. The continuous feed constant calculation procedures are summarized as follows:
\begin{enumerate}
	\item The simulation starts without injecting refueling materials (i.e. only removing FPs and Pa).
	\item After the first depletion calculation step, we check the total mass density of FPs and Pa in the \texttt{FP-waste tank} and \texttt{Pa-decay tank}, respectively.
	\item A simple calculation yields the amount of heavy metal that must be added during this cycle:
	\subitem 3a. \textbf{Thorium feed mechanism}: mass of injected $^{233}$U 
	$\approx$ mass of extracted Pa and mass of injected $^{232}$Th $\approx$ 
	mass of extracted FPs to maintain the total fuel mass constant.
	\subitem 3b. \textbf{Non-thorium feed mechanism}: mass of injected $^{233}$U $\approx$ mass of extracted Pa and mass of injected fuel (i.e., HALEU, Pu mixed with HALEU, reactor-grade Pu, or TRU) $\approx$ mass of extracted FPs to maintain the total fuel mass constant.
	\item Dividing this mass by time and inventory of refueling material gives the corresponding feed constant (1/s).
\end{enumerate}
In both feed mechanisms (thorium and non-thorium) we continuously added 
$^{233}$U and the mass of injected $^{233}$U is equal to the mass of extracted 
Pa ($^{233}$Pa decays to $^{233}$U after $\approx$ 27 d). To keep the total 
fuel mass almost constant, for \textbf{thorium feed mechanism}, the mass of 
injected thorium must be equal to the mass of extracted FPs and for 
\textbf{non-thorium feed mechanism}, the mass of injected fuel (i.e., HALEU, 
Pu mixed with HALEU, reactor-grade Pu, or TRU) must be equal to the mass of 
extracted FPs.

### 2020-02-25 paper/results.tex

For the Pu+HALEU case, $k_{eff}$ firstly increases during the first $\approx$ 
$3$ EFPY and then gradually decreases. This is due to the conversion of 
$^{238}$U from HALEU into fissile plutonium, which often produces the majority 
of the neutrons until the transition to Th/$^{233}$U fuel cycle. 
Figure~\ref{fig:diffPuAll811} shows the variation in fissile plutonium 
isotopes ($^{239}$Pu , $^{241}$Pu, and $^{243}$Pu) inventory during reactor 
operation. For TRU and reactor-grade Pu, the fissile plutonium isotopes are 
gradually depleted, during operating (the fissile material is dominated by Pu 
in these cases). For the HALEU case, the fissile material is dominated by 
$^{235}$U, however, $^{238}$U is converted into fissile plutonium. Therefore, 
the mass of fissile plutonium increases during operating but its value still 
relatively low. For the Pu+HALEU case, the mass of the fissile plutonium 
isotopes 
increases within the first $\approx$ $3$ EFPY and then decreases. The 
reactivity from the produced Pu is much greater than the depletion of the 
initial reactivity (due to the depletion of initial fissile isotopes), 
consequently, $k_{eff}$ increases during the first $\approx$ $3$ EFPY and 
then decreases; the amount of $^{233}$U generated in the Pu+HALEU case is 
insufficient to maintain the reactor criticality and counteract neutron 
absorption in non-fissile isotopes.

>>>>

For the Pu+HALEU case, $k_{eff}$ firstly increases during the first $\approx$ 
$3$ EFPY and then gradually decreases. This is due to the conversion of 
$^{238}$U from HALEU into fissile plutonium, which often produces the majority 
of the neutrons until the transition to Th/$^{233}$U fuel cycle. 
Figure~\ref{fig:diffPuAll811} shows the variation in fissile plutonium 
isotopes ($^{239}$Pu, $^{241}$Pu, and $^{243}$Pu) inventory during reactor 
operation. For TRU and reactor-grade Pu, the fissile plutonium isotopes are 
gradually depleted, during operating (the fissile material is dominated by Pu 
in these cases). For the HALEU case, the fissile material is dominated by 
$^{235}$U, however, $^{238}$U is converted into fissile plutonium. Therefore, 
the mass of fissile plutonium increases during operating but its value still 
relatively low. For the Pu+HALEU case, the mass of the fissile plutonium 
isotopes 
increases within the first $\approx$ $3$ EFPY and then decreases. The 
reactivity from the produced Pu is much greater than the depletion of the 
initial reactivity (due to the depletion of initial fissile isotopes), 
consequently, $k_{eff}$ increases during the first $\approx$ $3$ EFPY and 
then decreases; the amount of $^{233}$U generated in the Pu+HALEU case is 
insufficient to maintain the reactor criticality and counteract neutron 
absorption in non-fissile isotopes.
